A Real-Time Sonar Simulator for Automatic Target Recognition

### 2018-03-06 Simulator/article.tex
All compiling, but lots of work to be done

\let\oldincludegraphics\includegraphics
\renewcommand\includegraphics[2][]{%
  \immediate\write18{./bin/laFigure #2 #1}%
  \input{result}}%

% \DeclareMathOperator*{\argmin}{\text{arg}\;\text{min}}

\newcommand\Fig[1]{Fig.~\ref{#1}}
\newcommand\eq[1]{(\ref{#1})}

\newcommand\Grey[1]{{\color{Grey}#1}}
\newcommand\Red[1]{{\color{Red}#1}}
\newcommand\Blue[1]{{\color{Blue}#1}}
\newcommand\DarkBlue[1]{{\color{DarkBlue}#1}}
\newcommand\LightBlue[1]{{\color{LightBlue}#1}}
\newcommand\Brown[1]{{\color{Brown}#1}}
\newcommand\Green[1]{{\color{Green}#1}}
\newcommand\SeaGreen[1]{{\color{SeaGreen}#1}}
\newcommand\Yellow[1]{{\color{yellow}#1}}
\newcommand\Orange[1]{{\color{orange}#1}}

\newcommand\nn{\nonumber\\}

\newcommand\nmat[1]{\begin{matrix}#1\end{matrix}}
\newcommand\bmat[1]{\begin{bmatrix}#1\end{bmatrix}}
\newcommand\case[1]{\begin{cases}#1\end{cases}}
\newcommand\textbox[2]{\footnotesize\text{\parbox{#1}{\centering\emph{#2}}}}

\newcommand\rand{\text{rand}}
\newcommand\randn{\text{randn}}
\newcommand\rect{\text{rect}}
\newcommand\sinc{\text{sinc}}
\newcommand\tr{\text{tr}}
\newcommand\adj{\text{adj}}

% \newcommand\max{\text{max}}
\newcommand\argmin[1]{\text{arg}\;\underset{#1}{\text{min}}}

\newcommand\qqquad{\quad\qquad}
\newcommand\qqqquad{\qquad\qquad}

% \renewcommand\l[1]{\left#1}
% \renewcommand\r[1]{\right#1}

% {\text{\parbox{1.5cm}{\centering volume hyper- sphere}}}

%Keyword colouring:
\newcommand\kw[1]{#1}
\newcommand\parm[1]{#1}%\color{Black}#1\color{Black}}

\newcommand\of[1]{\scriptstyle(\parm{#1})\displaystyle}
\newcommand\df[1]{\scriptstyle[\parm{#1}]\displaystyle}
\newcommand\var[3]{#1_\text{#2}\of{#3}}

\newcommand\diag{\text{diag}}

% \raisebox{lift}[extend-above-baseline][extend-below-baseline]{text}
\newcommand\mt[1]{\text{\emph{#1}}} %mt = mathtext
\newcommand\mathnorm{\textstyle}
\newcommand\mathbig[1]{\displaystyle#1\mathnorm}
\newcommand\mathsmall[1]{\scriptstyle#1\mathnorm}
\newcommand\mathtiny[1]{\scriptscriptstyle#1\mathnorm}
\newcommand\sfrac[2]{\scriptstyle\raisebox{0.25pt}[0pt][0pt]{$\frac{#1}{#2}$}\mathnorm}
\newcommand\nfrac[2]{\textstyle\frac{#1}{#2}\displaystyle}

\newcommand\sumu[1]{\sum\limits^{#1}\;}
\newcommand\suml[1]{\sum\limits_{#1}\;}
\newcommand\sumb[2]{\sum\limits_{#1}^{#2}\;}

\newcommand\produ[1]{\prod\limits^{#1}\;}
\newcommand\prodl[1]{\prod\limits_{#1}\;}
\newcommand\prodb[2]{\prod\limits_{#1}^{#2}\;}

\newcommand\defeq{\overset{\underset{\mathrm{def}}{}}{=}}

%Math macros:
\newcommand\T{^{\scriptscriptstyle T}}
\renewcommand\H{^{\scriptscriptstyle H}}

\renewcommand\vec[1]{\boldsymbol{#1}}
\newcommand\mat[1]{\boldsymbol{#1}}

\newcommand\Om{O_\text{m}}
\newcommand\Oa{O_\text{a}}
\newcommand\Nl{N_\text{l}}
\newcommand\Nk{N_\text{k}}
\newcommand\1{\vec 1}
\newcommand\I{\mat I}
\renewcommand*\a{\vec a}
\newcommand*\f{\vec f}
\renewcommand*\i{\vec i}
\renewcommand*\k{\vec k}
\newcommand*\n{\vec n}
\newcommand*\p{\vec p}
\newcommand*\s{\vec s}
\newcommand*\w{\vec w}
\newcommand*\x{\vec x}
\newcommand*\y{\vec y}

\newcommand*\A{\mat A}
\newcommand*\B{\mat B}
% \newcommand*\C{\mat C}
\newcommand*\E{\mat E}
\renewcommand*\P{\mat P}
\newcommand*\eP{\mat{\hat P}}
\newcommand*\R{\mat R}
\newcommand*\Ri{\R^{-1}}
\newcommand*\eR{\mat{\hat R}}
\newcommand*\eRi{\hat{\mat R}\;\!^{-1}}
\newcommand*\Navg{N_\text{avg}}
\newcommand*\W{\mat W}
\newcommand*\X{\mat X}
\newcommand*\Xd{\X_{\!\Delta}}
\newcommand*\Y{\mat Y}

\renewcommand\argmin{\text{argmin}}

\renewcommand*\P{\mat P}
\newcommand*\V{\mat V}
\newcommand*\M{\mat M}

\renewcommand*\L{\mat \Lambda}
% \newcommand*\U{\mat U}
% \renewcommand*\t{\mathtiny{^T}}
% \newcommand*\h{\mathtiny{^H}}
\renewcommand*\t{^T}
\newcommand*\h{^H}

\newcommand\minus{\scalebox{0.75}[1.0]{$-$}}

% \usepackage{tikz}
% \usetikzlibrary{shapes,snakes}
\usepackage{amsmath,amssymb}
% \usepackage{datatool}
% \usepackage{glossaries}

% \newenvironment{outline}
% {\begin{itemize}}
% {\end{itemize}}

@@

\title{A Real-Time Sonar Simulator for Automatic Target Recognition}

\author{{Jo~Inge~Buskenes,~\IEEEmembership{Student Member,~IEEE,} %
        Herman~Midelfart, %
        \O{}ivind~Midtgaard} %
\thanks{J. I. Buskenes is with the Department of Informatics, University of Oslo, Norway.}%
\thanks{H. Midelfart and \O. Midtgaard are with The Norwegian Defence Research Establishment (FFI), Norway.}% <-this % stops a space
% \thanks{Manuscript received April 19, 2005; revised January 11, 2007.}
}

% The paper headers
\markboth{IEEE Journal of Oceanic Engineering}%
{A GPU Sonar Simulator for Automatic Target Recognition}

% Publishers ID mark:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
Template matching is a common technique for automatic classification of objects in synthetic aperture sonar (SAS) images. The principle is to isolate an image segment containing the object of interest, correlate it with a set of template images, and assign it to the class of the template yielding the highest correlation coefficient. The challenge is to come up with a representative set of template images covering the relevant configurations of object and seabed.

We target this challenge with a sonar simulator that first takes as input a seabed model derived from the real sonar image. Then it places a 3D object model on the seabed, renders the scene, and adds the resulting image to the template set. For every object type, position, alignment and material, the procedure is repeated, and a correlation coefficient computed. The best performance is obtained when these parameters are estimated from the sonar image as part of the classification process. The simulator is therefore written in OpenGL and OpenCL and runs on graphics processing units (GPUs). The result is a fast performing and portable template generator which can adapt to the characteristics of the current scene on-the-fly.
\end{abstract}

% Keywords (normally not used for peer reviews)
\ifPeerReview\else
\begin{IEEEkeywords}
Sonar, SAS, simulator, template matching, OpenGL, OpenCL.
\end{IEEEkeywords}
\fi}
% \fi

% make the title area
\maketitle

% This command fixes abstract positioning for compsoc articles:
\IEEEdisplaynotcompsoctitleabstractindextext

% (Optional) Add some extra info on cover page of peer review papers:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi

% Insert page break and insert second title (peer review mode)
\IEEEpeerreviewmaketitle

\section{Introduction}

\setcounter{totalnumber}{1}
\setcounter{topnumber}{1}
\setcounter{dbltopnumber}{1}

\begin{figure*}[t]\centering%
\includegraphics[width=\linewidth]{gfx/scene_coordinate_system.svg}%
\caption{Simulator principle. A model of the seafloor and object is first loaded into OpenGL, and an orthogonal camera view set up to match the desired image region on the seafloor. Then the scene is rendered and an intensity image and a depth map is created. The intensity image along with the depth information is finally combined to form a sonar like image. This is our template.}\label{simulator_coordinate_system}%
\end{figure*}

\begin{figure*}[t]\centering%
\includegraphics[drawing,width=\linewidth]{gfx/simulator.svg}%
\caption{Simulator concept.}\label{buildup}%
\end{figure*}

% \IEEEPARstart{W}{hen} 
% Autonomous underwater vehicles (AUVs). When an AUV with a sonar imaging system encounter an interesting object, it would be very beneficial if adapted its behavior to examine the object further. This is challenging as it would need to learn in what ways the interesting objects differ from the background, i.e. it relies on previous knowledge. It would also need to process this information is near real-time to act in an efficient manner.

% - Adapt simulation to its behaviour, see if they still match.

\IEEEPARstart{A}{utonomous} underwater vehicles (AUVs) are becoming increasingly important in human underwater endeavors. When equipped with a synthetic aperture sonar (SAS), they can e.g. be used to map the seafloor, survey underwater biology, help localize wrecks, inspect oil pipes, fields and electricity lines, or hunt for mines and biohazard waste. Such independent systems have much to gain from being able to adapt to the scenario at hand. For instance, with a well designed automatic target recognition (ATR) system on board they can adapt their mission plan to collect more information on particularly interesting objects, or fine-tune the sonar and navigation system to improve imaging performance.

% Automatic target recognition (ATR) is an important component in autonomous underwater vehicles (AUVs), as it allows the vehicle to adapt its mission plan to e.g. revisit detected objects for closer examination. One way to classify these objects is by using template matching. The principle involved  is to isolate an image segment containing the object of interest, compare it with a set of template images of the relevant object classes, and assign it to the class of the template with the the best fit.

In conventional ATR an object of interest is first isolated, its properties attempted estimated, and a classifier used to determine what object class these properties belong to. The challenge is finding unique and intrinsic object properties, such as its type, shape and size, and to do so independently of extrinsic properties such as its alignment, burial depth, degeneration level or view angle. While SAS provide images with state of the art resolution and dynamic range, they still only represent an ambiguous acoustical 2D projection of the 3D object, littered with a highly coherent noise floor.

An alternative approach is to use template matching, where the isolated image segment is compared to a set of template images covering the relevant object classes, and assigned to the class of the template with the best fit. Since the templates are computed from 3D models, this reduces the issue of image ambiguity. However, a challenge is finding a set of templates that represent the actual configurations of object and seafloor, Another issue is that the size of precomputed library of templates grows exponentially with the number of parameters needed. This limits its use to only a few parameters being coarsely sampled, ultimately yielding inaccurate results~\cite{Midelfart2010}.

% We avoid the problems of conventional template matching by computing the templates on-the-fly as part of the classification process, thereby not needing a precomputed set of templates at all. 
%  that avoids this problem by running sufficiently fast to create templates adapted to the actual scene as part of the classification process.
% Classification process

We tackle the problems of conventional template matching by performing:
%
\begin{itemize}
\item \emph{Adaptive template matching}; where a template is selected based on object and seafloor parameters estimated from the SAS image~\cite{Midelfart2010}, and
\item \emph{Real-time template simulation}; where a template is created tailored to the estimated parameters, improving its accuracy and avoiding the need of a precomputed template library.
\end{itemize}
%

The methods presented herein will focus on the template simulator, but some details will also be given on adaptive template matching for completeness. More details on this technique can be found in~\cite{Midelfart2010}.

 This is achieved with the aid of the massive computing power in graphics processing units (GPUs) and optimized software libraries for scene rendering. The simulator loads a 3D model of the seafloor and object class into OpenGL and positions the camera at the sonar location. The scene is highlighted using a light tube that extends to infinity along the AUV's axis of probagation. We assume rough, isotropic surfaces reflecting sound energy equally in all directions. This can be modeled with a Lambertian scattering model~\cite{Blake1993,Bell1995}, where the intensity of the backscattered sound solely depends on the incidence angle of the transmitted signal onto the model surface. The OpenGL rendering produces an optical 2D intensity image and depth map that reveals the distance from each pixel to the propagation axis. For maximum flexibility this data is finally combined with OpenCL to produce the image templates.

as part of the classification process. This is achieved with the aid of the massive computing power in graphics processing units (GPUs) and optimized software libraries for scene rendering.

The simulator loads a 3D model of the seafloor and an object class model into OpenGL, where emitted sound waves are modeled with a cylindrical light source placed at the sonar transmit location.

We assume rough, isotropic surfaces reflecting sound energy equally in all directions. This can be modeled with a Lambertian scattering model~\cite{Blake1993,Bell1995}, where the intensity of the backscattered sound solely depends on the incidence angle of the transmitted signal onto the model surface. When rendering OpenGL is set up to produce an optical 2D intensity image and depth map that reveals the distance from each pixel to the propagation axis. For maximum flexibility this data is finally combined with OpenCL to produce the image templates.

Various simulators for high frequency, side-looking sonar imagery have been published, e.g.~\cite{Bell1997,Sammelm2003}. However, most implementations have prioritized accurate acoustic modeling at the expense of execution speed. One exception is the SIGMAS+ simulator in~\cite{Coiras2009a, Coiras2009b}. Similarly to our approach it takes advantage of parallel processing on GPUs, and it uses OpenGL to render a 2D image from a 3D model, under the assumption of Lambertian scattering. However, SIGMAS+ is still aimed more towards realistically looking images, including effects like ambient noise, which is irrelevant for our purpose of template generation. Also, their simulator superimpose images obtained from multiple OpenGL rendering passes to create the sonar image, while we render once and post-process this result with OpenCL to tailor the sonar image with greater flexibility.

For lack of a better name, the simulator will be referred to as "FFISim" throughout this article.

Show:
\begin{itemize}
\item Adaptive template matching to avoid large static template libraries,
\item Improved classification performance by feeding simulator with parameters derived off SAS image,
\item Further improved classification by performing a fine-search within a narrow search space around the SAS derived parameters. GPU makes this more powerful.
\end{itemize}

\section{Methods}

Creating a 2D view of an arbitrary complex 3D scene is a non-trivial matter.  This is why we decided to use the core OpenGL pipeline to do this for us. OpenGL is a popular, well-matured and multi-platform application programming interface (API) for rendering 2D and 3D vector graphics. It relieves us from the intricacies of projecting vertices, faces and textures defined in a 3D space onto a suitable 2D image plane. This section explains how we set up OpenGL for this task, and then proceeds to describe how we post process the OpenGL images with OpenCL to form the final sonar image.

\begin{itemize}
\item Load arbitrary complex models
\item Simulate sonar with OpenGL
\item Speed considerations
\end{itemize}

Models are not ROCCAN (???)

\subsection{Loading the scene}

Our models are stored in regular 3D model files. The simulator begins by loading these into a tree of nodes, each representing a vertex in the model with corresponding facet and texture information. Then the vertices are placed into a virtual scene in OpenGL, and a camera set up to capture the part of the scene corresponding to the image. When OpenGL renders the image, it first computes which vertex contributes to which pixel in the image, then computes the intensity and color of the pixel based on the node's facet, texture and scattering model. The code that operates on vertices are called a vertex shader, and the code that operates on pixels are called a fragment shader.

The mapping of the vertices from object coordinates to image coordinates happen through a set of transformations:
%
\begin{align*}
\bmat{x\\y\\0\\1}_\text{image} = \boldsymbol{P} \cdot \boldsymbol{V} \cdot \boldsymbol{M} \cdot \bmat{x\\y\\z\\1}_\text{object},
\end{align*}
%
where $\bmat{\cdot}_\text{model}$ is the loaded model, $\bmat{\cdot}_\text{image}$ is the output image, and $\M$, $\V$ and $\P$ are transformation matrices. These will be described next.

The first transform we apply is the model matrix $\M$, which operates in the model's local coordinate system. It is defined as
%
\begin{align}
\M = \bmat{x_t \\ y_t \\ z_t \\ w}_\text{\parbox{1cm}{\setlength\baselineskip{0.3cm}transformed\\model}} &= \boldsymbol{T} \cdot \boldsymbol{R} \cdot \boldsymbol{S} \bmat{x \\ y \\ z \\ w}_\text{model},
\end{align}
%
where $\boldsymbol{S}$, $\boldsymbol{R}$ and $\boldsymbol{T}$ are matrices\todo{be explicit on dependencies?} for scaling, rotation and translation, respectively. These are provided for reference in Appendix \ref{transformation_matrices}.

Next we map the model into world coordinates using the view matrix $\V$. This involves aligning the camera to precisely fit the image region. In our case the camera represents the phase center of the sonar, and it is placed in the world origin. The transformation then take the form of
%
\begin{align}
\V &= \mat{T}(x_\text{c}, y_\text{c}, z_\text{c}) \cdot \R_\text{x}(\phi_x, 0, 0) \nonumber\\
&= 
\bmat{
1  &  0  &  0  &  x_\text{c} \\
0  &  1  &  0  &  y_\text{c} \\
0  &  0  &  1  &  z_\text{c} \\
0  &  0  &  0  &  1 \\
} \cdot
\bmat{
1  &  0           &  0           &  0 \\
0  &  \cos\phi_x  &  -\sin\phi_x &  0 \\
0  &  \sin\phi_x  &  \cos\phi_x  &  0 \\
0  &  0           &  0           &  1 \\
},
\end{align}
%
where $(x_\text{c}, y_\text{c}, z_\text{c})$ is the image center coordinates for along track, cross track, and depth, respectively, and $\phi_x$ is the elevation angle\todo{is this correct?} of the sonar. It is given as
%
\begin{align*}
\phi_x &= \arctan\left(\frac{y_\text{mean}}{-z_\text{mean}}\right).
\end{align*}\todo{use absolute sign?}

Finally the relevant part of the scene is projected onto the image plane with a projection matrix $\P$. There are two main types of projection, perspective and orthographic. Perspective projection scales all $x$ and $y$ coordinates inversely with the distance to the objects to create a sense of perspective in the image. Orthographic projection, on the other hand, maps all the scene's vertices to a squared cube where each dimension has the range [-1,1]. This will render all objects equally large whether they are in the distance or up close. The matrix that gives orthographic projection is defined as
%
\begin{align}
\P = \left(\begin{matrix}
\frac{2}{r-l}  &  0              &  0              &  -\frac{r+l}{r-l} \\
0              &  \frac{2}{t-b}  &  0              &  -\frac{t+b}{t-b} \\
0              &  0              &  \frac{-2}{f-n} &  -\frac{f+n}{f-n} \\
0              &  0              &  0              &  1
\end{matrix}\right)
\end{align}
%
where $l,r,b$ and $t$ are the left, right, bottom and top boundaries of the projection plane, respectively, and where $n$ and $f$ is the near and far limits of the cube that encompass the scene. These parameters need to be adapted to provide a 2D image that matches the image size, which we specify in along track and cross track coordinates ($x,y$). This is straightforward for left and right since matches the $x$-axis
% They define the window through which we observe our scene.
\begin{align}
l &= x_\text{min}  &  r &= x_\text{max}
\end{align}
%
For the near and far limits one simply needs to make sure the scene will fit in it, say
%
\begin{align}
n &= 0\;\text{m}   &  f &= 300\;\text{m}.
\end{align}
%
The bottom and top boundaries can be inferred from \Fig{camera_top_bottom}. First we define an origin vector that points to the center of the image
%
\begin{align}
\vec o &= \left[\frac{x_\text{min}+x_\text{max}}{2}, \frac{y_\text{min}+y_\text{max}}{2}, \frac{z_\text{min}+z_\text{max}}{2} \right]\T,
\end{align}
%
and the distance from the image origin to its $y$-boundaries is given as
%
\begin{align}
\Delta y &= \frac{|y_\text{max} - y_\text{min}|}{2}.
\end{align}
%
To find the nearest and farthest point in the image that will be part of the projected scene, $\alpha$ and $\beta$, we must consider the seafloor rotation angle around the $x$-axis, $\Phi_x$ 
%
\begin{align}
\vec\alpha &= \vec 0 - \left[0, \Delta y, \Delta y \cdot \tan\Phi_\text{x} \right] \nn
\vec\beta &= \vec 0 + \left[0, \Delta y, \Delta y \cdot \tan\Phi_\text{x} \right]
\end{align}

\begin{figure}[t]\centering%
\includegraphics[drawing,width=\linewidth]{gfx/opengl_part.svg}%
\caption{Setting camera top and bottom boundaries.}\label{camera_top_bottom}%
\end{figure}

\subsection{OpenGL rendering}

To produce the sonar templates we assume a rough, isotropic surface that reflects energy equally in all directions. This permits us to use a Lambertian scattering model where the backscatter intensity depends only on the incidence angle~\cite{Zhang1999}. It does not consider observation angle or sound frequency, but for the purpose of creating templates this is not needed.

The rendered image will appear as if we placed a window at the sonar and looked through it in the direction of the image. This window is resized to make sure that the template image perfectly fills it.

\subsection{OpenCL post-processing}

The ``camera image'' rendered with OpenGL is not in along-track and cross-track coordinates as we want it to be. It does, however, show us what parts of the scene that are visible from the sonar. OpenGL can also produce a depth map that reveals the distance to each of image pixels. This information can be converted to a ranged sonar-like image by simply adding up all the intensity values that share the same depth for each range line. We perform this computation in OpenCL. It allows general purpose programming on GPUs and can interoperate with OpenGL quite nicely. This way we keep all the calculations on the GPU.

\todo{
- Kernel tuning.
- Sharing. PBO. Asynchronous.
}

\subsection{Template matching}

1. Segmentation
2. Parameter estimation
   Available: Range, altitude, resolution => Input to both standard and adaptive. In standard the best fit out of a predefined template database is selected, then resampled to correct resolution.
   Additional for adaptive, parameters that modify object model: Target aspect angle, seafloor slope, object length and height.

   Target aspect angle computed with Radon Transform, seafloor under object modeled as plane with a slope estimated from bathymetric data, length estimated from longest part of highlight region, and height estimated from the shape of the shadow. The burial depth is n

3.

   Details: \ref{Midelfart2010}.

Let $S(x,y)$ be the isolated part of the

Matching the isolated image segment with the template take the form of

\begin{align}
c_{h,i} = \underset{x',y'}{\text{argmax}} \sum\limits_{\forall x}\sum\limits_{\forall y} \big| H(x,y) - T(x-x',y-y') \big|
\end{align}

\begin{align}
c_i = \frac{c_{h,i} + c_{s,i}}{2}
\end{align}

And the final template used is the one with the greatest correlation coefficient

\begin{align}
\underset{T_i(x,y)}{\text{argmax}} C_i\sum\limits_{\forall x}\sum\limits_{\forall y}  S(x,y) T(x-x',y-y')
\end{align}

Then assign object to class $i$.

To narrow down the search space of possible templates we derive some reliable statistics of the seafloor and object from the SAS image. While FFISim accepts an arbitrarily complex seabed, we find that modeling it as a plane wave with the right tilt works well. The object is more difficult. For instance, we estimate its length and height from the shape of its shadow, but this fails if the object is notably immersed into the seabed at one end.

After the a template is created we decompose it into its highlight and shadow component, and compute a correlation score between these and the corresponding areas in the SAS image. Then the two scores are summed and normalized to form the final correlation score of the template.

We tag the template matchers as either semi- or fully adaptive. The former refers to using the best matching SIGMAS+ generated template out of a precomputed set of templates, while the latter refers to the proposed method of using FFISim to generate templates on the-the-fly. \todo{Should be more precise. How many templates? Parameter span? How finely were the parameters sampled?}

% It is also possible to compute the sonar image using the OpenGL ``blend'' feature.

\subsection{Speed considerations}

Using OpenGL and OpenGL to generate sonar templates on the GPU allow hundreds of sonar templates to be formed per second. While this is sufficient for our needs. However, initializing the OpenGL and OpenCL context and loading 3D models from file into it is a slow process, and typically takes several seconds for every invocation. Therefore, to alleviate the simulation performance we either had to supply it with a list of parameters or ensure that it did not have to be shut down for every invocation.

The rendering loop that uses OpenGL and OpenCL to form

\todo{
Write to file
}

%    """
%    
%    stm: Standard template matching, made with SIGMAS.
%    atm: Adaptive template matching, made with FFISim.
%    atm_bury: Adaptive template matching where a search of bury depths is used.
%    atm_bury_restr: Restricted?
%    
%    Adaptiveness made by letting the template matcher use 5 different rotation
%    angles in the ground-range axis. 
%    
%    
%    Data from Jesusbukta
%    
%    Figures are from run 2012.06.16
%    det210 - Cylinder 1: est 2.92m (short side towards sonar)
%    det233 - Cylinder 2: est 3.05m (diagonal)
%    det379 - Torpedo: est 5.06m (looks good)
%    
%    ROC curves from 3 different runs, first two just 2 days apart in 2009, last one in 2012.
%    
%    run_20120616 - last one. Much better with adaptive templates.
%    
%    ROC curves made by average correlation score of shadow/highlight from template matching.
%    
%    
%    fig4b(2015.11.16)
%       Bug that skipped masking of port side (negative y) detections past 180m range fixed
%       
%    fig4 (2015.11.16)
%       Bug leading to incorrect estimation of cylinder orientation fixed.
%          This affects run090601_1. Should be better after this.
%          
%    fig3 (2015.11.09)
%       Detections outside 180m range have been masked (but bug for port side fixed in fig4b).
%       Fixed an incorrectly annotated detection.
%       
%    fig2 (2015.11.02)
%       Height of object estimated from shadow length
%       Let adaptive method try 5 different rotations in ground-range (-4, -2, 0, 2, 4)
%          Improved results, but incorrect classification of:
%          5 detections in first run, and
%          3 detections in second run
%          => For FPR=0.2 the results are slightly worse than with normal adaptive (particularly first run).
%          Images are problematic in these cases:
%             2 detections get wrong orientation because a line (anchor) cross the image
%             Other misclassifications in image regions with poor echo or shadow strength
%          => Not a big issue as not enough time to detect more than ~20% of the detections in a run
%             In a MCM operation these detections would never be used.
%             The runs have 3000 detections each, photographing 20% of these is too much.
%          Performance FPR>0.2 not interesting.
%          Performance FPR<0.2 much better with adaptive template matching.
%          
%    fig (2015.10.23)
%       atm - cylinder orientation and length estimated from SAS image and fed into simulator.
%       
%       ROC curves from last run. Computing the mean correlation score of the shadow and highlight/echo.
%       Early results. Can only check vs. cylinder-like objects.
%       Standard template library created based on estimation of physical size of known cylinder mines.
%       Not possible for adaptive method to perform better when hitting such a target.
%       
%  
%    """

\section{Results \& Discussion}

\begin{figure*}[!tbp]\centering%
\includegraphics[width=\linewidth]{gfx/fig_images_sonar_simulator_tagged.svg}\label{fig_images_sonar_simulation}%
\caption{Selected objects. Top row show SAS images of 2 cylinders and a torpedo located in the Jesus Bay, Norway. Center row show the best match from a conventional template matcher, and bottom row show the adaptive template matcher. The contour of the SAS object is indicated with red lines. Note that the adaptive technique produce templates with a better fit to the highlight, but that this does not necessarily lead to a better match against the shadow.}
\end{figure*}

\begin{figure*}[!tbp]\centering%
\includegraphics[width=\linewidth]{gfx/fig_rocs.pdf}\label{fig_roc_curves}%
\caption{Receiver operating characteristics (ROC) curves.}
\end{figure*}

>>>>

\input{macros}

%\let\oldincludegraphics\includegraphics
%\renewcommand\includegraphics[2][]{%
%  \immediate\write18{./bin/laFigure #2 #1}%
%  \input{result}}%
%  
%% \DeclareMathOperator*{\argmin}{\text{arg}\;\text{min}}
%
%\newcommand\Fig[1]{Fig.~\ref{#1}}
%\newcommand\eq[1]{(\ref{#1})}
%
%\newcommand\Grey[1]{{\color{Grey}#1}}
%\newcommand\Red[1]{{\color{Red}#1}}
%\newcommand\Blue[1]{{\color{Blue}#1}}
%\newcommand\DarkBlue[1]{{\color{DarkBlue}#1}}
%\newcommand\LightBlue[1]{{\color{LightBlue}#1}}
%\newcommand\Brown[1]{{\color{Brown}#1}}
%\newcommand\Green[1]{{\color{Green}#1}}
%\newcommand\SeaGreen[1]{{\color{SeaGreen}#1}}
%\newcommand\Yellow[1]{{\color{yellow}#1}}
%\newcommand\Orange[1]{{\color{orange}#1}}
%
%\newcommand\nn{\nonumber\\}
%
%\newcommand\nmat[1]{\begin{matrix}#1\end{matrix}}
%\newcommand\bmat[1]{\begin{bmatrix}#1\end{bmatrix}}
%\newcommand\case[1]{\begin{cases}#1\end{cases}}
%\newcommand\textbox[2]{\footnotesize\text{\parbox{#1}{\centering\emph{#2}}}}
%
%\newcommand\rand{\text{rand}}
%\newcommand\randn{\text{randn}}
%\newcommand\rect{\text{rect}}
%\newcommand\sinc{\text{sinc}}
%\newcommand\tr{\text{tr}}
%\newcommand\adj{\text{adj}}
%
%% \newcommand\max{\text{max}}
%\newcommand\argmin[1]{\text{arg}\;\underset{#1}{\text{min}}}
%
%\newcommand\qqquad{\quad\qquad}
%\newcommand\qqqquad{\qquad\qquad}
%
%% \renewcommand\l[1]{\left#1}
%% \renewcommand\r[1]{\right#1}
%
%% {\text{\parbox{1.5cm}{\centering volume hyper- sphere}}}
%
%%Keyword colouring:
%\newcommand\kw[1]{#1}
%\newcommand\parm[1]{#1}%\color{Black}#1\color{Black}}
%
%\newcommand\of[1]{\scriptstyle(\parm{#1})\displaystyle}
%\newcommand\df[1]{\scriptstyle[\parm{#1}]\displaystyle}
%\newcommand\var[3]{#1_\text{#2}\of{#3}}
%
%\newcommand\diag{\text{diag}}
%
%% \raisebox{lift}[extend-above-baseline][extend-below-baseline]{text}
%\newcommand\mt[1]{\text{\emph{#1}}} %mt = mathtext
%\newcommand\mathnorm{\textstyle}
%\newcommand\mathbig[1]{\displaystyle#1\mathnorm}
%\newcommand\mathsmall[1]{\scriptstyle#1\mathnorm}
%\newcommand\mathtiny[1]{\scriptscriptstyle#1\mathnorm}
%\newcommand\sfrac[2]{\scriptstyle\raisebox{0.25pt}[0pt][0pt]{$\frac{#1}{#2}$}\mathnorm}
%\newcommand\nfrac[2]{\textstyle\frac{#1}{#2}\displaystyle}
%
%\newcommand\sumu[1]{\sum\limits^{#1}\;}
%\newcommand\suml[1]{\sum\limits_{#1}\;}
%\newcommand\sumb[2]{\sum\limits_{#1}^{#2}\;}
%
%\newcommand\produ[1]{\prod\limits^{#1}\;}
%\newcommand\prodl[1]{\prod\limits_{#1}\;}
%\newcommand\prodb[2]{\prod\limits_{#1}^{#2}\;}
%
%\newcommand\defeq{\overset{\underset{\mathrm{def}}{}}{=}}
%
%%Math macros:
%\newcommand\T{^{\scriptscriptstyle T}}
%\renewcommand\H{^{\scriptscriptstyle H}}
%
%\renewcommand\vec[1]{\boldsymbol{#1}}
%\newcommand\mat[1]{\boldsymbol{#1}}
%
%\newcommand\Om{O_\text{m}}
%\newcommand\Oa{O_\text{a}}
%\newcommand\Nl{N_\text{l}}
%\newcommand\Nk{N_\text{k}}
%\newcommand\1{\vec 1}
%\newcommand\I{\mat I}
%\renewcommand*\a{\vec a}
%\newcommand*\f{\vec f}
%\renewcommand*\i{\vec i}
%\renewcommand*\k{\vec k}
%\newcommand*\n{\vec n}
%\newcommand*\p{\vec p}
%\newcommand*\s{\vec s}
%\newcommand*\w{\vec w}
%\newcommand*\x{\vec x}
%\newcommand*\y{\vec y}
%
%\newcommand*\A{\mat A}
%\newcommand*\B{\mat B}
%% \newcommand*\C{\mat C}
%\newcommand*\E{\mat E}
%\renewcommand*\P{\mat P}
%\newcommand*\eP{\mat{\hat P}}
%\newcommand*\R{\mat R}
%\newcommand*\Ri{\R^{-1}}
%\newcommand*\eR{\mat{\hat R}}
%\newcommand*\eRi{\hat{\mat R}\;\!^{-1}}
%\newcommand*\Navg{N_\text{avg}}
%\newcommand*\W{\mat W}
%\newcommand*\X{\mat X}
%\newcommand*\Xd{\X_{\!\Delta}}
%\newcommand*\Y{\mat Y}
%
%\renewcommand\argmin{\text{argmin}}
%
%
%\renewcommand*\P{\mat P}
%\newcommand*\V{\mat V}
%\newcommand*\M{\mat M}
%
%\renewcommand*\L{\mat \Lambda}
%% \newcommand*\U{\mat U}
%% \renewcommand*\t{\mathtiny{^T}}
%% \newcommand*\h{\mathtiny{^H}}
%\renewcommand*\t{^T}
%\newcommand*\h{^H}
%
%\newcommand\minus{\scalebox{0.75}[1.0]{$-$}}
%
%% \usepackage{tikz}
%% \usetikzlibrary{shapes,snakes}
%\usepackage{amsmath,amssymb}
%% \usepackage{datatool}
%% \usepackage{glossaries}
%
%% \newenvironment{outline}
%% {\begin{itemize}}
%% {\end{itemize}}

@@

\input{document}
